---
title: "Ebola Hackathon"
author: "Will Townes"
date: "November 22, 2014"
output: html_document
---

```{r}
library(plyr)
library(lars)
```
Data Cleaning
```{r}
covariates<-read.csv('data/ebola_covariate_data.csv',header=TRUE)
#drop all columns with all NA
covariates<-Filter(function(x)!all(is.na(x)), covariates)
#table(covariates$sdr_level)

ts2<-read.csv('data/original/02.csv',header=TRUE)
id.map<-read.csv('data/original/01.csv',header=TRUE)

#religious
cov.relig<-read.csv('data/original/15.csv',header=TRUE)
head(cov.relig)
#problem- the SDR_IDs for relig are at lower geographic level (liberia only), need to convert to ADM1 level instead of ADM2
#cov.relig$sdr_id_adm1<-
lookup.relig<-unique(cov.relig[,"ADM1_NAME"])
id.map.relig<-id.map[id.map$level=="ADM1" & id.map$name %in% lookup.relig,c("name","level","id","parent_id")]
#Rivercess is called "River Cess" in the Covariates table but "Rivercess" in the religion table. From manual inspection, "River Cess" is correct with SDR_ID=77115
id.map.relig[id.map.relig$name=="Rivercess","id"]<-c(77115)
id.map.relig<-id.map.relig[1:(nrow(id.map.relig)-1),]
# change column names to match the religion table
id.map.relig<-id.map.relig[,c("name","id")]
colnames(id.map.relig)<-c("ADM1_NAME","parent_sdr_id")
cov.relig<-join(cov.relig,id.map.relig,by="ADM1_NAME")
relig.summary<-ddply(cov.relig,c("parent_sdr_id"),summarise,nChurch=sum(TYPE=="Church"),nMission=sum(TYPE=="Mission"),nMosque=sum(TYPE=="Mosque"),nTemple=sum(TYPE=="Temple"))
colnames(relig.summary)[1]<-"sdr_id"
#merge religious data into main covariates data
covariates<-join(covariates,relig.summary,by="sdr_id")
#remove old data frames to free up memory
rm(cov.relig,id.map.relig,relig.summary)
#remove columns with all NA
covariates<-Filter(function(x)!all(is.na(x)), covariates)
#remove extraneous columns
drops<-c("X","pos","country_code","iso_code","gdlcode","region","sdr_level")
covariates<-covariates[,!(colnames(covariates) %in% drops)]
covariates<-covariates[,c(31,32,1,2,3:30,33:36)]
#write out cleaned covariates data
#write.csv(covariates,"data/covariates2.csv")
#covariates<-read.csv("data/covariates2.csv",header=TRUE)
cor(covariates[,6:23])
#use edyr gender difference instead of separate variables, since both highly correlated with edyr
covariates<-within(covariates,edyr_gender_diff<-edyr_male-edyr_fem)
drops<-c("edyr_fem","edyr_male")
covariates<-within(covariates,age_under_20<-age09+age1019)
covariates<-within(covariates,age_over_60<-age6069+age7079+age90hi)
drops<-c(drops,"age09","age1019","age2029","age3039","age4049","age5059","age6069","age7079","age8089","age90hi")
covariates<-covariates[,!(colnames(covariates) %in% drops)]
covariates<-covariates[,-1] #drop "X" column
cor(covariates[,5:ncol(covariates)])
#small house highly inv. correlated with large house. Drop small house
drops<-c("small_house")
#floor quality highly corr with bad_floor. Drop bad_floor
drops<-c(drops,"bad_floor")
#bad_toilet correlated with flush_toilet (and subjective), drop bad_toilet
drops<-c(drops,"bad_toilet")
#iwi, tv, phone, electr, floor quality, and toilet all highly correlated. Expect we will use only one of these. Not clear which to drop though.
covariates<-covariates[,!(colnames(covariates) %in% drops)]
meta<-covariates[,c(1,2,3,4)] #split off non-numeric info
covariates<-covariates[,-c(2,3,4)]
#split off religion info
relig.cols<-c("nChurch","nMission","nMosque","nTemple")
cov.relig<-covariates[,c("sdr_id",relig.cols)]
covariates<-covariates[,!(colnames(covariates) %in% relig.cols)]
#at this point we have 16 potential covariates, and 4 religious covariates applicable only in Liberia

#fitted lambda(response of interest) and population/area data provided by Nick and Leah Comment
data.lambda<-read.csv("data/fitted_lambda.csv",header=TRUE)
#remove cases with NA response.
data.lambda<-data.lambda[complete.cases(data.lambda),]
#add population density column and log-transform lambdas
data.lambda<-within(data.lambda,log.lambda<-log(lambda))
data.lambda$Population<-as.numeric(as.character(data.lambda$Population))
data.lambda<-within(data.lambda,pop.density<-Population/Area_km2)
hist(data.lambda$pop.density)
#highly skewed, use log instead
data.lambda<-within(data.lambda,log.pop.dens<-log(pop.density))
#drop descriptive column
data.lambda<-data.lambda[,-3]
cor(data.lambda[,2:7])
#interestingly log.lambda is inv. correlated with population and area but not population density.
#include log of Area and log of Population
data.lambda<-within(data.lambda,log.pop<-log(Population))
data.lambda<-within(data.lambda,log.area<-log(Area_km2))
cor(data.lambda[,2:9])
#less collinearity between log pop,area versus pop, area, so use logs and drop original. Also removes skewness. Also remove pop.density columns since linear comb of pop and area on log scale
hist(data.lambda$log.pop)
hist(data.lambda$log.area)
data.lambda<-data.lambda[,-c(3,4,6,7)]
dat.full<-join(data.lambda,covariates,by="sdr_id")
```
Try lasso linear regression with response variables (lambdas- parameter for infectiousness) log-transformed.
```{r}
y<-as.vector(dat.full$log.lambda)
x<-as.matrix(dat.full[,4:ncol(dat.full)])
colnames(x)<-colnames(dat.full)[4:ncol(dat.full)]
cv.lars(x,y,K=10,type="lar",use.Gram=FALSE)
cv.lars(x,y,K=5,type="lar",use.Gram=FALSE)
model<-lars(x,y,type="lar",use.Gram=FALSE)
summary(model)
coef(model)[2,]
```
The LASSO based on minimizing cross-validation error chooses only the intercept model (5 or 10 fold CV). The LASSO based on Cp chooses the model with only the log.area term. This suggests 
```{r}
yhat<-predict(model,x,2,type="fit")$fit
resid<-y-yhat
plot(yhat,resid)
hist(resid)
which(resid==min(resid))
#observation 4 may be outlier
```
Try using principal components instead
```{r}
x.pc<-svd(x)$u[,1:3]
cv.lars(x.pc,y,K=10,type="lar",use.Gram=FALSE)
cv.lars(x.pc,y,K=5,type="lar",use.Gram=FALSE)
model<-lars(x.pc,y,type="lar",use.Gram=FALSE)
summary(model)
```
Principle Components also not significant